Depression and suicidality as costly signals of need

### 2021-04-01 Paper.rmd
added ecdf plot

tar_load(
  c(
    'd',
    'power_curve',
    'models',
    'models_prereg',
    'effect_sizes',
    'effect_sizes2',
    'effectsPlots', 
    'plot_T1Belief_T1Action',
    'plot_emotions',
    'plot_raw_beliefs',
    'plot_raw_actions',
    'med1b',
    'med4',
    'm1b_boot',
    'm3b_boot',
    'dfT1emotion_vignette',
    'dfT1emotion_signal',
    'dfT2emotion_signal'
    )
  )

@@

# Introduction
Since Durkheim’s pioneering effort to use quantitative social science to explain suicide (cite), anthropologists, psychologists, and sociologists have made varied attempts at explaining depression and suicide. Despite this effort and substantial investment by funding agencies, rates of depression and suicide attempts have remained largely unchanged over the last 100 years (confirm and cite), and there is still substantial uncertainty about why both are as common as they are (cite). Likely due to their high costs and the suffering they cause, the vast majority of these approaches have been built on the assumption that they are the result of internal dysfunction (cite examples). However, a subset of anthropologists and evolutionary theorists have challenged this idea, suggesting that despite the subjective suffering they cause, they may function as a means to limit negative fitness outcomes in times of substantial adversity (Firth + others).

@@

## Connection to Adversity
Likely the most reliable finding when it comes to depression and suicidal behavior is that most cases are closely tied to major negative life events (Hammen, 2005 + others; citations for suicide). Compared to non-depressed individuals, those with depression tend to report at least twice as many instances of negative events in both community and clinical samples (Mazure, 1998) and more negative events than those with schizophrenia and bipolar depression across a wide range of studies (Paykel, 1994). In instances where depression is preceded by negative events, the onset of depression tends to be rapid, often occurring within one month of the event (Kendler et al., 1999 [double check and find others]).

This strong connection between negative life events and depression along with the need and distress that likely accompanies these events has led to the suggestion that this relationship is causal (Kendler et al., 1999). Support for this comes from the fact that the relationship between negative events and depression holds true when looking solely at events outside of one's control, which shows that the connection is unlikely to be driven solely by depressed individuals creating situations where negative events are likely to be common (Hammen, 2005; Kendler et al., 1999). In addition to this, twin studies have shown that one's history of negative events is still a strong predictor of major depression when controlling for genetic similarity and that part of the heritability of depression stems from the heritability of negative events, like divorce and family conflict (Kendler & Baker, 2007; Kendler et al., 1999).

@@

Suicide attempts also tend to co-occur with social adversity (Hagen et al., 2008 + cite and describe), with evidence of this existing cross-culturally. Grief over the loss of a loved one (Firth; Devries et al. 2011), bullying (Alavi et al., 2017; Butler et al., 2019; Ford et al., 2017; Wang et al., 2020), and instances of social transgression (Firth; Malinowski; Syme and Hagen 2018; find Western citation to confirm) have all been reported to lead to suicidal behavior in both industrialized (confirm the transgression part) and traditional societies. Of the recorded instances of suicide attempts in the Human Relations Area Files (HRAF), 77.4% of attempts were found to be linked to times of social conflict, 88.7% were linked to fitness threats in general, and 84.9% were linked to situations in which attempters were powerless to resolve their situation alone (Syme et al., 2016). As with depression, assault is an particularly powerful risk factor for suicidal behavior based on multiple large-scale multi-country studies (Devries et al. 2011; Ellsberg et al. 2008) and more localized examinations. For example, both sexual and non-sexual assault were also associated with more self-harm and suicidal behaviors among both indigenous Sami and majority Norwegian adolescents, with those who attempted suicides also reporting more than twice the number of negative life events than those who did not (Reigstad and Kvernmo 2017).

@@

Beyond resulting in the cost differential that allows for costly signaling in the first place, the costs of depression and suicidal behavior may be particularly important in times of conflict or when the level of support needed to address one’s situation is substantial. Importantly, there are likely few forms of social adversity that do not come with conflict. For example, breakups in both romantic and cooperative relationships are often preceded by conflict (Amato & Hohmann-Marriott, 2007), and surely involve conflict when one side wants reconciliation and the other does not. Likewise, the death of a loved one may also be a source of conflict as the survivor may turn to others for more support who might not want to invest more in the survivor (Hagen, 2003). Regardless of the source of conflict, greater conflict should correspond with greater conflicts of interests, something which would favor greater investment in manipulating others through fake signals. In these instances, less costly signals like verbal requests, facial sadness, and crying are likely to be unreliable as their low costs can easily be made up for with increased social support, thereby favoring the use of costlier signals in those experiencing need they cannot resolve on their own.

In addition to signaling need to those invested in the well-being of the signaler, depression and suicidal behavior may also help an individual bargain for increased support. This is possible due to the reduced investment in others that frequently accompanies depression or the threat to future cooperation that comes from a suicide. In these instances, depression and suicide attempts are suggested to function analogously to a labor strike, sending the message that one’s need is genuine and that if the target of the signal wanted to continue receiving benefits from the depressed, he or she would need to increase support towards the signaler.

@@

Beneficial social changes are also reported in traditional societies. Unlike many Western countries, where suicide is often viewed as pathological (Hidaka, 2012), members of traditional societies have been reported to view suicide attempts as cries for help rather than mental illness (Shostak, 1981). For example, attempts have been described as ways of escaping unwanted marriage arrangements, persistent abuse, or a lack of support in obtaining mates by those engaging in the behavior and observers (Gutierrez de Pineda & Muirden, 1948; Hilger, 1957; Karsten, 1935; Tessmann, 1930; Wilson, 1960), suggesting that help is at least occasionally provided or reliably expected. In support of this, 30 out of 84 examples of suicidal behavior included in the HRAF resulted in positive changes for the survivor, even though suicide was rarely if ever the main topic (Syme et al., 2016).

@@

## notes: 
add funding info somewhere
update MTurk (Ed email links + mine)
Find a place for the predictions and preregistration 
go back and emphasize conflict and PI in the methods (at least in the pilot studies portion)
add abstract
add conclusion

@@

This study involved a between-subjects pretest-posttest design to examine how different emotional signals would influence (1) the degree participants believed a fictional character to be in need and (2) the likelihood they would provide help within a vignette study.

@@

Four vignettes were used in this study in which both the presence of conflicts of interest and the existence of private information about one's need were emphasized. Due to the expectation that depression and suicide are likely too costly of signals for relatively minor instances of need, we designed this study to involve severe scenarios associated with suicide attempts both ethnographically and clinically (i.e, sexual assault (cite), non-sexual assault (cite), and thwarted marriage (cite)).

Participants in the US sample were randomly assigned to either the "basketball coach," "romantic partner," or "brother-in-law" vignettes. In the "basketball coach" scenario, the star player on a university basketball team comes to the participant (who is asked to imagine themselves as the athletic director at a university) and claims she was sexually assaulted by the coach. However, there are indications the player might be lying to get more playing time.

The "brother-in-law" vignette involved participants imagining that they let their sister, brother-in-law, and niece move in with the participant and their 15-year-old daughter after their sister's family lost their house in a fire. During this time, the participant's daughter becomes jealous of her cousin before accusing her uncle of sexually assaulting her.

The "romantic partner" vignette had participants imagine that they found a highly desirable partner after years of being single. However, their 13-year-old daughter, who has a history of trying to get in the way of their relationships is clearly not happy with this. After a period of sustained conflict, she accuses the romantic partner of physically assaulting her but cannot produce any evidence.

Those in the Indian sample only saw the "thwarted marriage vignette." In this scenario, the participant and their family is described as trying to arrange a dowry for their older daughter so she can marry the man she loves. After the man's family demands more money, the participant's family tries to find a second man, who the older daughter claims to find unattractive. Due to the younger daughter in the story only being two years younger, any additional investment in the dowry will come at the younger daughter's expsense.

The full vignettes are available in the SI.

@@

As the amount of crying in those with depression is moderately associated with depression severity [@VingerhoetsEtAl2016] (check this. IIRC he might have a more recent paper saying it's not associated with depression as people tend to think) and depression is a major risk factor for suicide [@BostwickPankratz2000; @Kessler2012], each signal was designed to build off of the less costly signals so that each signal contained elements of any signal it was costlier than. The full text for the signals are availible in the SI.

@@

Participants for this study were recruited from Amazon Mechanical Turk, a crowdsourcing platform that allows for the creation of Human Intelligence Tasks (HITs) that workers can complete for pay. As Amazon provides the infrastructure, it allows for a relatively low-cost way of collecting data for academic research at the cost of the data not being representative of any real population (Thomas and Clifford 2017). Despite this limitation, it is more diverse than most convenience samples [@KennedyEtAl2018(2020 now); @DworkinEtAl2016; (Thomas and Clifford 2017)], with turkers in the US mainly differing in being younger, more educated, and lower income than the greater U.S. population [@RossEtAl2010]; Boas et al. 2018). Indian samples are also likely to be younger and more educated than the general population but differ from US samples in that they are less diverse, have higher incomes relative to other Indians, and are more likely to come from regions with good internet access (Boas et al. 2018)

Overall, the quality of data provided by MTurk workers tends to resemble that of university sample pools[@NeckaEtAl2016, Robinson et al. 2019], with there being some indication that it might outperform university samples due to its previously mentioned diversity (Thomas and Clifford 2017) or the overall attentiveness of turkers [@HauserSchwarz2016]. For this reason, concerns about data quality come primarily from the threat of bot use or respondents faking their location to take surveys in a language they do not understand well, with there being little evidence of the former [@KennedyEtAl2018] and the risk of the latter able to be minimized through well designed attention checks, timed responses, or general study design (Aguinis et al. 2020; Huang et al. 2015; Kennedy et al. 2020; Thomas and Clifford 2017).

@@

To reduce the risk of low quality responses, we used a variety of attention checks, which we clearly labeled in an attempt to avoid demand characteristics. The first was shown immediately after the consent form and provided participants with a random word and asked them to enter the vowels in the order in which they are found in the word. Capitalization did not matter and no word contained repeat vowels. Failure to enter the correct vowels in order resulted in the participants being filtered out of the study. The second followed the vignette and involved asking three questions about the story that should be easy to answer for anyone paying attention.

We also filtered out those who read the vignette faster than one-third the time it took MG to read it as fast as possible without concern for comprehension. This occured immediately after participants finished the vignette.

Informed consent was given by proceeding with the study after reading the consent form, and all participants who passed the attention checks were paid $1 for their time, with us estimating it would take 4-8 minutes before data collection. This study was certified exempt by the Washington State University Human Research Protection Program (certification number: 18528-001).

@@

After accepting the HIT on MTurk, every participant took the survey using Qualtrics. In every instance, the elements of the survey were laid out in the following order: 1) the consent form, 2) the vowel attention check, 3) the main body of the vignette, 3) the pre-test baseline measures (T1), 4) the story content attention check, 5)  the signal portion of the vignette, 7) the post-test outcome measures (T2-T3), 8) an attention check that did not filter out participants (T4), 9) a set of demographic questions, and 10) an open feedback prompt. The full survey can be found in the SI.

@@

In both samples, we recorded self-reports on how much participants believed the signaler to be telling the truth (T1Belief) and the likelihood of them helping the signaler as requested (T1Action). With the thwarted marriage vignette we also asked how they would split the money they had saved for the dowry between their daughters (T1Divide). In every instance, the order of the questions was randomized to avoid order effects [@KrosnickAlwin1987].

Responses were recorded with sliders due to the fact they allow for finer grained changes than categorical scales [@KlimekEtAl2017]. In every instance the position started at the left due to findings that people may be less likely to move the slider from central starting positions (cite and word better). For the exact wording of the question and the labels on the sliders, see table 1.

@@

As both a validity check and a way to understand the degree participants would be willing to help if they believed the participant completely, we presented participants with strong evidence that the claims were true (T3Action). In the US sample, this involved video evidence of the event in question occurring or a similar event after the fact. With the thwarted marriage vignette, this involved the participant seeing the man their daughter wants to marry trash talk their daughter and their family. As a final attention check, we presented participants with the T3Action question but included instructions to ignore this and drag the slider all the way to the left (T4AttentionCheck). We did not exclude anyone from the study for failing this nor did we withhold payment.

@@

1,950 participants started the study by clicking the link to Qualtrics (1,213 American and 737 Indian). After removing participants who did not finish the survey, failed attention checks, or moved through the study at an unrealistc pace, our sample dropped to N = 1,240 (937 American and 303 Indian). Basic summary statistics on the remaining N = 1,240 participants (Male = 759, Female = 479) can be found in table XX.

@@

## Distributions of beliefs and actions at time 1

Power analyses prior to data collection showed that high proportions of participants scoring extremely high on belief or likelihood of helping at T1 would result in low power for this study's design, a problem which we appeared to largely avoid (see figure XX).

Although not a part of our pre-registered predictions, both T1Belief and T1Action were highly correlated across the four vignettes (enter correlation coefficients), consistent with help being worth providing if the signaler's claims were true (see figure XX).

@@

As a whole, our predictions for the effect of signal cost on belief and likelihood of helping T2 were met, with costlier signals eliciting greater pro-signaler responses. There was also support for an interaction between signal and T1Belief or T1Action, although the effect was not large (See figure 2). However, the effect of signal cost varied by vignette, with signal type having the greatest effect in the "basketball coach" and "brother-in-law" vignettes. When ordered according to cost, the greatest jump in support tended to occur between crying and mild-depression, with the next largest jump tending to occur between depression and suicide attempt. Overall, there was little difference between the effect of mild depression and the effect of depression.

```{r beliefeffect, fig.width=8, fig.height=8, fig.cap="The effect of the signals on T2 Belief (top) and Action (bottom) controlling for T1 values of Belief and Action, respectively, with an interaction between T1 values and the signal. Estimated using generalized linear regression models with the quasibinomial family."}
(effectsPlots$m1bplot + theme(axis.text.x = element_blank()))/effectsPlots$m3bplot + plot_layout(ncol = 1) & theme(plot.margin = unit(c(0,0,0,0), units = 'mm'))
```

```{r beliefvignette, fig.width=8, fig.height=8, fig.cap=paste0("The effect of the signals on T2 Belief (top) and Action (bottom) in each vignette, controlling for T1 values of Belief and Action, respectively. Effects computed with T1 values set to their median value (T1Belief=", median(d$T1Belief, na.rm=T), ", T1Action=", median(d$T1Action, na.rm=T), "). Estimated using generalized linear regression models with the quasibinomial family.")}
(effectsPlots$m2plot + theme(axis.text.x = element_blank()))/effectsPlots$m4plot + plot_layout(ncol = 1) & theme(plot.margin = unit(c(0,0,0,0), units = 'mm'))
```

@@

The total effect of the *Depression* signal (treatment) vs. *Verbal request* (control) on the likelihood of *Action* was to increase it by `r 100*signif(med1b$tau.coef, 3)` points from T1 to T2 (on the original 100-point scale). Of this increase, `r 100*signif(med1b$n.avg, 2)`% was mediated by the increased *Belief* that the victim was telling the truth. See Figure \@ref(fig:mediation). The total effect of the *Suicide attempt* signal (treatment) vs. *Verbal request* (control) on the likelihood of *Action* was to increase it by `r 100*signif(med4b$tau.coef, 3)` points from T1 to T2. Of this increase, `r 100*signif(med4b$n.avg, 2)`% was mediated by the increased *Belief* that the victim was telling the truth (Check if med4 or med4b is correct; they look identical to me).

@@

We also examined the effect of signaling on how participants perceived the signaler's emotional state. To do this, we created four new variables: LowMood (T1 and T2) and Manipulative (T1 and T2). This was done by adding the dummy variables for whether a participant thought the signaler was depressed, distressed or sad for the LowMood variables and by adding the dummy variables for whether they thought the particpant was Devious or Jealous for the Manipulative variables. In all instances, 0 indicated a participant did not select the emotion, while 1 indicated selection.

We then plotted the mean change seen in the data in addition to the change seen in 500 bootstrap resamples (see figure 6).

@@

Consistent with costly signaling theory (is this a common enough prediction to say this?), costlier signals resulted in greater perceptions of need and increased the likelihood of participants helping the signaler than the less costly signals. This effect was largest in the "brother-in-law" and "romantic partner" vignettes, both of which involved claims of assault against participants' imagined daughters. However, the effect of signal type was much smaller in both the "basketball coach" and the "thwarted marriage" vignette. In the "basketball coach" vignette, part of the reason for this may have been the lack of relatedness between participants and the alleged victim. However, this was still a surprising finding due to the relatively small size of the request (i.e., recommend the coach be suspended rather than kick out family or a romantic partner). With the thwarted marriage vignette, the fact that support came at the cost of one's younger daughter may explain the relatively small effect of signal strength. However, data from the Indian sample appears to be of lower quality, limiting our confidence in this interpretation (see limitations for more information).

In the "brother-in-law" and "romantic partner" scenarios, there was a noticeable jump in the magnitude of pro-signaler responses when going from the crying to mild depression and then a second, smaller jump from depression to a suicide attempt. One potential reason for this is the increase in symptoms from mild depression to depression was not costly enough to result in a large effect as the symptoms only differed in severity (e.g., grades dropping from As to Bs [mild depression] or Cs [depression]). Such small changes may be less impactful in vignette form than in real-life scenarios in which the effects of signaling may increase in severity and be prolonged over a lengthy period of time.

We also found evidence for the predicted interaction between signal and belief in need and likelihood of help at T1, with the pro-signaler effects being larger at intermediate but low levels of believe at T1. In real-life scenarios, it would be reasonable to expect that costly-signaling would not be engaged in when others already believe one's need. For this reason, too strong of a signal may be seen as manipulative or a cue of dysfunction, and therefore, be responded to aversively. However, belief in need and likelihood of help at T2 still remained high when it was high at T1, suggesting that this may not be a significant risk when one is in severe enough need.

@@

Although this study was not designed to examine whether the costs of signaling would be made up for by the increased support stemming from depression and suicide attempts directly, the results are encouraging (reword). In general, the costs required to honestly convey one's qualities should be expected to increase with the potential payoffs of deception (cite), with the vignettes in this study emphasizing the existence of private information and sizable conflicts of interest to increase the perceived risk of being taken advantage of. In similar real-world scenarios, this difficulty of signaling honest need would coincide with the substantial fitness-relevant benefits of preventing future assaults or being able to marry the partner of one's choice. These benefits coupled with the fact (too strong?) that depression and suicide attempts become relatively less costly with increased need (Hagen 2003), suggest that the increased support seen in this study may be worth the costs required to signal need more often than is often appreciated when counting the entire costs of depression as the signal rather than only the upregulation of symptoms for a communicative purpose.

@@

We estimated power using our pilot study data by comparing the effect of the *Depression* signal vs. the *Verbal request* signal on *Belief* (that the sister needed the money) at T2 controlling for *Belief* at T1 in a linear regression model. Specifically, we resampled the pilot data with replacement with sample sizes ranging from N=20 to N=100, fit the linear regression model, and extracted the p-value for the signal coefficient, repeating this 1000 times for each sample size. Power was computed as the proportion of p-values < 0.05. Power was about 80% for a sample size of about 35, and was about 90% for a sample size of about 45.

>>>>

tar_load(
  c(
    'd',
    'power_curve',
    'power_curve2',
    'models',
    'models_prereg',
    'effect_sizes',
    'effect_sizes2',
    'effectsPlots', 
    'ecdf_plot',
    'plot_T1Belief_T1Action',
    'plot_emotions',
    'plot_raw_beliefs',
    'plot_raw_actions',
    'plot_raw_belief_action2d',
    'med1b',
    'med4',
    'm1b_boot',
    'm3b_boot',
    'dfT1emotion_vignette',
    'dfT1emotion_signal',
    'dfT2emotion_signal'
    )
  )

@@

# Introduction
Since Durkheim’s pioneering effort to use quantitative social science to explain suicide, varied attempts by anthropologists, psychologists, and sociologists and substantial investment by funding agencies, rates of depression and suicide attempts have remained largely unchanged over the last 100 years (confirm and cite), and there is still substantial uncertainty about why both are as common as they are (cite). Due to their high costs, the vast majority of these approaches have been built on the assumption that they are the result of internal dysfunction (cite examples). However, a subset of anthropologists and evolutionary theorists have challenged this idea, suggesting that despite the subjective suffering they cause, they may function as a means to limit negative fitness outcomes in times of substantial adversity (Firth + others).

@@

## Connection to Adversity
Likely the most reliable finding when it comes to depression and suicidal behavior is that most cases are closely tied to major negative life events (Hammen, 2005 + others; citations for suicide). Compared to non-depressed individuals, those with depression tend to report at least twice as many instances of negative events in both community and clinical samples (Mazure, 1998) and more negative events than those with schizophrenia and bipolar depression across a wide range of studies (Paykel, 1994). In instances where this is the case, the onset of depression tends to be rapid, often occurring within one month of the event (Kendler et al., 1999 [double check and find others]).

This strong connection between negative life events and depression along with the need and distress that likely accompanies these events has led to the suggestion that this relationship is causal (Kendler et al., 1999). Support for this comes from the fact that the relationship between negative events and depression holds true when looking solely at events outside of one's control, which shows that the connection is unlikely to be driven solely by depressed individuals creating situations where negative events are likely to be common (Hammen, 2005; Kendler et al., 1999). In addition to this, twin studies have shown that one's history of negative events is still a strong predictor when controlling for genetic similarity and that part of the heritability of depression stems from the heritability of negative events, like divorce and family conflict (Kendler & Baker, 2007; Kendler et al., 1999).

@@

Suicide attempts also tend to co-occur with social adversity (Hagen et al., 2008 + cite and describe), with evidence of this existing cross-culturally. Greif over the loss of a loved one (Firth; Devries et al. 2011), bullying (Alavi et al., 2017; Butler et al., 2019; Ford et al., 2017; Wang et al., 2020), and instances of social transgression (Firth; Malinowski; Syme and Hagen 2018; find Western citation to confirm) have all been reported to lead to suicidal behavior in both industrialized (confirm the transgression part) and traditional societies. As with depression, assault is a powerful risk factor for suicidal behavior, something which has been found in multiple large-scale multi-country studies (Devries et al. 2011; Ellsberg et al. 2008). Both sexual and non-sexual assault were also associated with more self-harm and suicidal behaviors among both indigenous Sami and majority Norwegian adolescents, with those who attempted suicides reporting more than twice the number of negative life events than those who did not (Reigstad and Kvernmo 2017). The connection between social adversity and suicide is also seen in the Human Relations Area Files (HRAF), in which 77.4% of attempts were found to be linked to times of social conflict, 88.7% were linked to fitness threats in general, and 84.9% were linked to situations in which attempters were powerless to resolve their situation alone (Syme et al., 2016).

@@

Beyond resulting in the cost differential that allows for costly signaling in the first place, the costs of depression and suicidal behavior may be particularly important in times of conflict or when the level of support needed to address one’s situation is substantial. Importantly, there are likely few forms of adversity that do not come with conflict. For example, breakups in both romantic and cooperative relationships are often preceded by conflict (Amato & Hohmann-Marriott, 2007), and surely involve conflict when one side wants reconciliation and the other does not. Likewise, the death of a loved one may also be a source of conflict as the survivor may turn to others for more support who might not want to invest more in the survivor (Hagen, 2003). Regardless of the source of conflict, greater conflict should correspond with greater conflicts of interests, something which would favor greater investment in manipulating others through fake signals. In these instances, less costly signals like verbal requests, facial sadness, and crying are likely to be unreliable as their low costs can easily be made up for with increased social support, thereby favoring the use of costlier signals in those experiencing need they cannot resolve on their own. Importantly, the amount of conflict

In addition to signaling need to those invested in the well-being of the signaler, depression and suicidal behavior may also help an individual bargain for increased support. This is possible due to the reduced investment in others that frequently accompanies depression or the threat to future cooperation that comes from a suicide. In these instances, depression is suggested to function analogously to a labor strike, sending the message that one’s need is genuine and that if the target of the signal wanted to continue receiving benefits from the depressed, he or she would need to increase support towards the signaler.

@@

Beneficial social changes are also reported in traditional societies. Unlike Western countries, where suicide is often viewed as pathological (Hidaka, 2012), members of traditional societies have been reported to view suicide attempts as cries for help rather than mental illness (Shostak, 1981). For example, attempts have been described as ways of escaping unwanted marriage arrangements, persistent abuse, or a lack of support in obtaining mates by those engaging in the behavior and observers (Gutierrez de Pineda & Muirden, 1948; Hilger, 1957; Karsten, 1935; Tessmann, 1930; Wilson, 1960), suggesting that help is at least occasionally provided or reliably expected. In support of this, 30 out of 84 examples of suicidal behavior included in the HRAF resulted in positive changes for the survivor, even though suicide was rarely if ever the main topic (Syme et al., 2016).

@@

## notes: 
add funding info somewhere
update MTurk (Ed email links + mine)
Find a place for the predictions and preregistration 
go back and emohasize conflit and PI in the methods (at least in the pilot studies portion)

@@

This study involved a vignette study with between subjects pre-test/post-test design with one factor (signal) and 5 levels (verbal request, crying, crying with behavioral changes, depression, and suicide attempt), with the outcome of interest being how believe and likelihood of helping changed in response to a given signal.

## Pilot Studies

The first pilot study used a vignette study to examine how the presence of conflict, private information, and emotional signaling impacted the degree participants believed the signaler to be in need and provide support using an MTurk sample of 1,447 Americans. Of the signals, depression resulted in the most pro-signaler responses (reword), separating itself from the less costly signals when conflict and private information was present and substantially outperforming the schizophrenic break control. However, the suicide attempt condition was responded to relatively poorly, contrary to predictions.

One potential explanation is that the initial scenario of a mother asking their sibling for help paying for their young daughter's heart surgery was not severe enough for a suicide attempt to be a functional signal, due to a lack of beleivability or due to its costs appearing manipulative. For this reason, we designed this study with more severe scenarios that were also associated with suicide attempts both ethnographically and clinically, including instances of sexual assasult, non-sexual assault, and thwarted marriage.

add that we ask now ask same question twice? (interpretabilty)

The second was posted on the samplesize subreddit with the goals of testing the beleivability of the new vignettes and how pro-signaler participants were at T1, with a power analysis with simulated data showing this would limit our power. In this small sample of 28, we saw no evidence that beleivabilty was a problem and no evidence that participants as a whole would be too pro-sister at T1.

@@

Four vignettes were used in this study, with participants in the US sample being randomly assigned to either the "basketball coach," "romantic partner," or "brother-in-law" vignettes. In the "basketball coach" scenario, the star player on a unversity basketball team comes to the participant (who is asked to imagine themselves as the athletic director at a university) and claims she was sexually assaulted by the coach. However, there are indications the player might be lying to get more playing time.

The "brother-in-law" vignette involved participants imagining that they let their sister, brother-in-law, and niece move in with the participant and an imagined 15-year-old daughter after they lost their house in a fire. During this time, the participant's daughter becomes jealous of your niece before accusing her uncle of sexually assaulting her.

The "romantic partner" vignette hard participants imagine that they found the best partner they have ever had after years of being single. However, their 13-year-old daughter, who has a history of trying to get in the way of their relationships is clearly not happy with this. At one point, she accuses the romantic partner of physically assaulting her but cannot produce any evidence.

Those in the Indian sample only saw the "thwarted marriage vignette." In this scenario, the participant and their family is trying to arrange a dowry for their older daughter so she can marry the man she loves. After his family demands more money, the participant's family tries to find another man, who the older daughter claims to find unattractive. However, it is unclear how unattractive he is and any additional investment in the dowry will come at the expense of the younger daughter.

The full vignettes are available in the supplementary information.

@@

As the amount of crying in those with depression is moderately associated with depression severity [@VingerhoetsEtAl2016] and depression is a major risk factor for suicide [@BostwickPankratz2000; @Kessler2012], each signal was designed to build off of the less costly signals so that each signal contained elements of any signal it was costlier than.

@@

Participants for this study were recruited from Amazon Mechanical Turk, a crowdsourcing platform that allows for the creation of Human Intelligence Tasks (HITs) that workers can complete for pay. As Amazon provides the infrastructure, it allows for a relatively low-cost way of collecting data for academic research at the cost of the data not being representative of any real population. Despite this limitation, it is more diverse than most convenience samples [@KennedyEtAl2018; @DworkinEtAl2016], with turkers in the U.S. mainly differing in being younger, more educated, lower income, and more female biased than the greater U.S. population [@RossEtAl2010].

Overall, the quality of data provided by MTurk workers tends to resemble that of university sample pools, with there being some indication that it might outperform university samples in a few important areas. This is seen in that while both MTurk workers and university students engage in problematic behaviors that might lessen data quality to similar amounts [@NeckaEtAl2016], workers on MTurk have been shown to be more attentive than samples of university students [@HauserSchwarz2016]. For this reason, concerns about data quality come primarily from potential bot use and respondents faking their location to take surveys in a language they do not understand well, with the former concern being largely unfounded and the latter able to be corrected for with the tools used in this study [@KennedyEtAl2018]

@@

To ensure participants were paying attention, we used a variety of attention checks, which we clearly labeled in an attempt to avoid demand characteristics. The first was shown immediately after the consent form and provided participants with a random word and asked them to enter the vowels in the order in which they are found in the word. Capitalization did not matter and no word contained repeat vowels. Failure to enter the correct vowels in order resulted in the participants being filtered out of the study. The second followed the vignette and involved asking three questions about the story that should be easy to answer for anyone paying attention.

We also filtered out those who read the vignette faster than one-third the time it took MG to read it as fast as possible without concern for comprehension immediately after the vignette.

Informed consent was given by proceeding with the study after reading the consent form, and all participants who passed the attention checks were paid $1 for their time. (mention time here)

@@

After accepting the HIT on MTurk, every participant took the survey using Qualtrics. In every instance, the elements of the survey were laid out in the following order: 1) the consent form, 2) the vowel attention check, 3) the main body of the vignette, 3) the pre-test baseline measures (T1), 4) the story content attention check, 5)  the signal portion of the vignette, 7) the post-test outcome measures (T2), 8) an attention check that did not filter out participants, 9) a set of demographic questions, and 10) an open feedback prompt. The full survey can be found in the SI.

@@

In both samples, we recorded self-reports on how much participants believed the signaler to be telling the truth (BeliefT1) and the likelihood of them helping the signaler as requestedc (ActionT1). With the thwarted marriage vignette we also asked how they would split the money they had saved for the dowry between their daughters (Divide T1). In every instance, the order of the questions was randomized to avoid order effects [@KrosnickAlwin1987].

Responses were recorded with sliders due to the fact they allow for finer grained changes than categorical scales [@KlimekEtAl2017]. In every instance the position started at the left due to findings that people may be less likely to move the slider when it starts in the center (cite and word better). For the exact wording of the question and the ends of each continuum, see table 1.

@@

As both a validity check and a way to understand the degree participants would be willing to help if they believed the participant completely, we presented participants with strong evidence that the claims were true (Action T3). In the US sample, this involved video evidence of the event in question occuring or a similar event after the fact. With the thwarted marriage vignette, this involved the participant seeing the man their daughter wants to marry trash talk their daughter and their family. As a final attention check, we presented participants with the Action T3 question but included instructions to ignore this and drag the slider all the way to the left (T4AttentionCheck). We did not exclude anyone from the study for failing this nor did we withhold payment.

@@

@@

## Distributions of beliefs and actions at Time 1

@@

```{r beliefeffect, fig.width=8, fig.height=8, fig.cap="The effect of the signals on T2 Belief (top) and Action (bottom) controlling for T1 values of Belief and Action, respectively, with an interaction between T1 values and the signal. Fit using generalized linear regression models with the quasibinomial family."}
(effectsPlots$m1bplot + theme(axis.text.x = element_blank()))/effectsPlots$m3bplot + plot_layout(ncol = 1) & theme(plot.margin = unit(c(0,0,0,0), units = 'mm'))
```

```{r beliefvignette, fig.width=8, fig.height=8, fig.cap=paste0("The effect of the signals on T2 Belief (top) and Action (bottom) in each vignette, controlling for T1 values of Belief and Action, respectively. Effects computed with T1 values set to their median value (T1Belief=", median(d$T1Belief, na.rm=T), ", T1Action=", median(d$T1Action, na.rm=T), "). Fit using generalized linear regression models with the quasibinomial family.")}
(effectsPlots$m2plot + theme(axis.text.x = element_blank()))/effectsPlots$m4plot + plot_layout(ncol = 1) & theme(plot.margin = unit(c(0,0,0,0), units = 'mm'))
```

```{r ecdf, fig.width=11, fig.height=6, fig.cap="Empirical cummulative distribution functions for Time 2 Belief (top) and Action (bottom) compared to their Time 1 baseline values (red), by signal and vignette. Dots indicate median values of Belief and Action, by signal and vignette."}
ecdf_plot
```

@@

The total effect of the *Depression* signal (treatment) vs. *Verbal request* (control) on the likelihood of *Action* was to increase it by `r 100*signif(med1b$tau.coef, 3)` points from T1 to T2 (on the original 100-point scale). Of this increase, `r 100*signif(med1b$n.avg, 2)`% was mediated by the increased *Belief* that the victim was telling the truth. See Figure \@ref(fig:mediation).

@@

@@

Consistent with costly signaling theory, costlier signals resulted in greater perceptions of need and increased the likelihood of participants helping the signaler than the less costly signals. This effect was largest in the "brother-in-law" and "romantic partner" vignettes, both of which involved claims of assault against participants' imagined daughters. However, the effect of signal type was much smaller in both the "basketball coach" and the "thwarted marriage" vignette. In the "basketball coach" vignette, part of the reason for this may have been the lack of relatedness between participants and the alleged victim. However, this was still a surprising finding due to the relatively small size of the request (i.e., recommend the coach be suspended rather than kick out family or a romantic partner). With the thwarted marriage vignette, the fact that support came at the cost of one's younger daughter may explain the relatively small effect of signal strength. However, data from the Indian sample appears to be of lower quality, limiting our confidence in this interpretation (see limitations for more information).

In the "brother-in-law" and "romantic partner" scenarios, there was a noticeable jump in the magnitude of pro-signaler responses when going from the crying to mild depression and then a second, smaller jump from depression to a suicide attempt. One potential reason for this is the increase in symptoms from mild depression to depression was not costly enough to result in a large effect as the symptoms only differed in severity (e.g., grades dropping from As to Bs [mild depression] or Cs [depression]). Such small changes may be less impactful in vignette form than in real-life scenarios in which the effects of signaling may increase in severity and be prolonged over a lenghty period of time.

We also found evidence for the predicted interaction between signal and belief in need and likelihood of help at T1, with the pro-signaler effects being larger at intermediate but low levels of believe at T1. In real-life scenarios, it would be reasonable to expect that costly-signaling would not be engaged in when others already believe one's need. For this reason, too strong of a signal may be seen as manipulative or a cue of dysfunction, and therefore, be responded to aversively. However, belief in need and likelihood of help at T2 still remained high when it was high at T1, suggesting that this may not be a significant risk when one is in severe enough need (want to bring in MA here. Is that a bad idea?).

@@

Although this study was not designed to examine whether the costs of signaling would be made up for by the increased support stemming from depression and suicide attempts directly, the results are encouraging. In general, the costs required to honestly convey one's qualities should be expected to increase with the potential payoffs of deception (cite), with the vignettes in this study emphasizing the existence of private information and sizable conflicts of interest to increase the perceived risk of being taken advantage of. In similar real-world scenarios, this difficulty of signaling honest need would coincide with the substantial fitness-relevant benefits of preventing future assaults or being able to marry the partner of one's choice. These benefits coupled with the fact (too strong?) that depression and suicide attempts become relatively less costly with increased need (Hagen 2003), suggest that the increased support seen in this study may be worth the costs required to signal need more often than is often appreciated when counting the entire costs of depression as the signal rather than only the upregulation of symptoms for a communicative purpose.

@@

```{r}
pwr80_90 <- 
  power_curve %>% 
  group_by(Base) %>% 
  summarise(
    pwr80 = min(sample_size[power>=.8]), 
    pwr90 = min(sample_size[power>=.9])
  )
```
